---
title: "ESM 204 Assignment 3"
author: "Julia Wilson & Thomas Lenihan"
output: html_document
date: '2022-05-06'
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, options(scipen = 999))

library(tidyverse)
library(here)
library(janitor)
library(thematic)
library(ggplot2)
library(equatiomatic)

```


## Overview

```{r, include = FALSE}
# Read in data

electricity <- read_csv(here("data", "hw3_data.csv")) %>% 
  clean_names() %>% 
  select(-x1)

```

The Biden Administration’s “interim” value is $51 per metric ton of CO2. The electricity sector is the second largest source of greenhouse gas emissions in the U.S. (after transportation). In this analysis, we will consider the distributional consequences of imposing a household electricity tax based on the SCC to address the climate change problem.

## Analysis

The data includes the following: 
- Consumers can be separated into two income groups: “high” and “low.” 
- Price (in $) and quantity (in kWh) estimates of demand per month for the two groups. 

The analysis will show linear regressions (with an intercept) to estimate the demand curves for “high” and “low” income consumers under the following scenarios: 

-  Initially, there is no tax on electricity consumption.
- The current electricity price (without any taxes) is $.10 per kWh.
- The marginal cost of producing a kWh of electricity is linear and has a price-intercept of 0.

### **1. Marginal External Cost of electricity**

Assuming one kWh of electricity emits 0.85 pounds of CO2 and the interim SCC is $51/ton, we can then calculate the MEC per kWh. 

```{r, include = FALSE}

# Create a longer data set switching income from a column to a variable

electricity_long <- electricity %>% 
  pivot_longer(cols = c(q_low_kwh, q_high_kwh),
               names_to = 'income_level',
               values_to = 'kwh') %>%
  mutate(income_level = case_when(income_level == 'q_low_kwh' ~ 'low',
                                  income_level == 'q_high_kwh' ~ 'high'))

# Run a calculation to get the MEC 

# Set up the interim price at $51/ton
interim_price_ton <- 51

# Convert price per ton to price/lb 
interim_price_lb <- interim_price_ton/2205

# Multiply the interim price by 0.85 to get MEC/kWh 
interim_price_kWh <- interim_price_lb * 0.85

```

The marginal external cost per kWh of electricity is 1.97 cents.

### **2. Aggregations**

 

**Low Income Demand Curve:**
```{r}
# lm estimates for both low and high incomes 
demand_low <- lm(price_cents ~ kwh, income_level == 'low',
                 data = electricity_long)
#extract_eq(model = demand_low, use_coefs = TRUE, coef_digist = 8)

#demand_low
```
The equation for our low income demand curve was **P = 23.37 - 0.00011Q**

**High Income Demand Curve:**
```{r}
demand_high <- lm(price_cents ~ kwh, income_level == 'high',
                  data = electricity_long)
#extract_eq(model = demand_high, use_coefs = TRUE, coef_digist = 5)

#demand_high
```

The equation for our high income demand curve was **P = 31.61 - 0.000052Q**


```{r}
# Demand model from sample code

demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
```


```{r}
# Aggregate demand from code sample 

demand_agg <- function(p){
  q <- demand(p, demand_low) + demand(p, demand_high)
  return(q)
}

# demand_agg(10) = 536,719.5
```

With the income-based demand curves, we can sum horizontally to find the **aggregate demand curve:**
```{r}
# Make a vector and extract the lm 
price = seq(0, 30, length.out = 100)

Qagg <- map(price, demand_agg) %>% 
  unlist()

agg_df<- tibble(Qagg = Qagg, price = price)

demand_agg_eq <- lm(price ~ Qagg, data = agg_df) 
```

```{r}
# Switch the equations above so Q is on one side and then add

agg_slope <- demand_agg_eq$coefficients[2]
agg_int <- demand_agg_eq$coefficients[1]

#agg_slope
#agg_int

# Aggregate demand function is P = 29.78135 - 0.00004(Qagg)

```

```{r, include = FALSE}

# [OLD CODE, CAN DELETE]

# Convert the equations so that Q is on one side and the sum

demand_low_flip <- lm(kwh ~ price_cents, income_level == 'low', data = electricity_long)

demand_high_flip <- lm(kwh ~ price_cents, income_level == 'high', data = electricity_long)

kwh_agg_int <- (demand_low_flip$coefficients[1]+demand_high_flip$coefficients[1])

kwh_agg_slope <- (demand_low_flip$coefficients[2] +
                    demand_high_flip$coefficients[2])

demand_agg_p <- function(price_cents){kwh_agg_int + kwh_agg_slope*(price_cents)}

demand_agg_int <- kwh_agg_int*-kwh_agg_slope
demand_agg_slope <- 1/kwh_agg_slope

# Aggregate demand function: P = 30.5 - 0.000039Q

demand_agg <- function(kwh){demand_agg_int + demand_agg_slope *kwh}
  
```

The resulting aggregate demand equation from 0 to 158,642 is: 

**P = 31.61 - 0.000052Q**

and from 158,462 to 820,339, it is: 

**P = 29.78 - .00004Q**

(The price is in cents, not dollars)

**Supply Curve**

The current price is $0.10/kWh and the MC for electricity is linear with a 0 intercept. 

```{r}
kwh_agg <- demand_agg_p(10)

# 536,719.5 kWh at $0.10

supply_slope <- 10/kwh_agg
# 0.000019
```

The supply curve will pass through the agg demand curve at $0.10/kwh. We can then infer the supply curve equation will be: 

**P = 0.0000192Q**

#### Visualizing the data

```{r}
ggplot(agg_df, aes(Qagg, price)) +
  geom_line(color = "goldenrod3") +
  annotate("text", x = 400000, y = 17, 
           label = "Demand (Agg)", angle = -30) +
  geom_abline(color = "firebrick4",
              intercept = demand_high$coefficients[1],
              slope= demand_high$coefficients[2]) +
  annotate("text", x = 300000, y = 15, 
           label = "High-Income Demand", angle = -38) +
  geom_abline(color = "steelblue4",
              intercept = demand_low$coefficients[1],
              slope= demand_low$coefficients[2]) +
  annotate("text", x = 150000, y = 9, 
           label = "Low-Income Demand", angle = -60) +
  geom_abline(color = "darkolivegreen",
              intercept = 0,
              slope= supply_slope) +
  annotate("text", x = 300000, y = 7, 
           label = "MPC", angle = 17) +
  labs(x = "Electricity Consumed (kWh)",
       y = "Price (cents/kWh)") +
  theme_minimal()
```

#### **Figure 1: Market for Electricity**



We can then calculate consumer and producer surplus and environmental cost. We will be cautious of converting from cents to dollars. 

```{r}
# Consumer Surplus 

# Converting cents to dollars for the y-axis
cs <- 0.5*kwh_agg*(0.305-0.10)
#cs
# $53,518.12

# Producer Surplus 

ps <- 0.5*kwh_agg*(.10)
# $26,106.39

# Environmental Cost 
env_cost <- interim_price_kWh*kwh_agg
# $10,264.96
```

**Calculating Consumer Benefit**: Take the area under the aggregate demand curve, but
above the current price of electricity ($0.10/kWh).This area is a triangle with base 536,719.5
and height 19.78.
0.5 * 536,719.5 * (29.78-10) = 5308155 cents

**Consumer Surplus = $53,081.5**

**Calculating Producer Benefit**: Take the area under the aggregate demand curve, but above
the current price of electricity ($0.10/kWh).This area is a triangle with base 536,719.5 and
height 10.
0.5 * 536,719.5 * 10 = 2683595 cents

**Producer Surplus = $26,835**

**Calculating Environmental Cost:** Multiply the MEC found in question 1 (0.019) by Q*
(536,719.5)
0.0197 * 536,719.5 =  $10,573.4

 **Environmental Cost = $10,573.4**

### **3. Consumer Benefit**

**Low income consumers:** consumer benefit is found by taking the area under the low income demand curve but above the current price of electricity ($0.10/kWh). This area is a triangle with base 121545 and height 13.37. 

  0.5 * 13.37 cents/kWh * 121545 kWh = 812,528 cents 
  
  **benefit for low income consumers is $8,125.3.**

**High income consumers:**consumer benefit is found by taking the area under the high income demand curve but above the current price of electricity ($0.10/kWh). This area is a triangle with base 415577 and height 21.61.

  0.5 * 21.61 cents/kWh * 415577 kWh = 4,490,309.5 cents

  **benefit for high income consumers is $44,903.1.**
  
total consumer surplus = 8,125.3 + 44,903.1 = $53,028.4

total consumer surplus = (13.37 * 158,641.5) + ((31.61 - 23.37) * 0.5  * 158,641.5) + (0.5 * (525,641  - 158,641.5) * 13.37)

total consumer surplus = **$52,280.3**

(8,125.3/53,028.4)*100 = 15.3%
(44,903.1/53,028.4)*100 = 84.7%

**low income consumers receive 15.3% of the total benefit.**

**high income consumers receive 84.7% of the total benefit. **



### **4. Optimal Tax**

```{r, include = FALSE}
mpc_q <- function(p) {p/supply_slope} # MPC in terms of Q
# DEMAND FUNCTION WITH TAX (SOLVED IN #1 MEC = $0.01965986 -> 1.97 cents)
dem_tax <- function(p, model){
  q <- (p - (model$coefficients[[1]]-1.97))/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
# AGG DEMAND FUNCTION w/ TAX
dem_agg_tax <- function(p){
  q <- dem_tax(p, demand_low) + dem_tax(p, demand_high)
  return(q)
}
# NEW EQUILIBRIUM 
uniroot(function(p)
  dem_agg_tax(p) - mpc_q(p),
        interval = c(0,20))
# a. Amount of electricity produced and consumed
# b. Price of electricity
#save new p
p_tax <- 9.31988
#generate new Q
q_tax <- dem_agg_tax(p_tax)

#c. Welfare of "high" income consumers
demand_high_tax <- dem_tax(p_tax, demand_high)
cs_high_tax <- (0.5 * (.3161 - 0.0197 - (p_tax/100)) * demand_high_tax) - (demand_high_tax * 0.0197)

#d. Welfare of "low" income consumers
cs_low_og <- cs - env_cost
demand_low_tax <- demand(p_tax, demand_low)
#Below is the total env cost at the new level of consumption

env_cost_tax <- interim_price_kWh * q_tax
cs_low_tax <- (0.5 * (.2337 - 0.0197 - 0.0932) * demand_low_tax) - env_cost_tax #- (demand_low_tax * 0.0197)

#cs_low_tax
# e. Power suppliers 
# FIND NEW P FOR SUPPLIERS GIVEN Q*
q_tax <- dem_agg_tax(p_tax)
ps_tax <- 0.5*q_tax*0.0932

# f. Total environmental damage
# P* of MEC (0.0197) TIMES Q*
env_change <- env_cost - env_cost_tax

# g. Tax revenue generated
# P* TIMES Q*
tax_rev <- interim_price_kWh*q_tax

q_tax_diff <- q_tax - kwh_agg

p_tax_consumer <- p_tax + (interim_price_kWh*100)
```


a) the new quantity of electricity produced and consumed with the tax will be approximately **`r round(q_tax, 2)`kWh,**  which is **`r round(q_tax_diff, 2)` kWh** less than the status quo. 

b) The new price of electricity with the tax will increase from 10 cents to **`r round(p_tax_consumer,2)` cents.**


c) welfare effect of tax on high income consumers: **$`r round(cs_high_tax, 2)`** NEED TO COMPARE TO STATUS QUO

d) welfare effect of tax on low income consumers:  **$`r round(cs_low_tax, 2)`**

e) overall effect of tax on power generators: a **$`r round(ps_tax, 2)`** decrease

f) The total environmental damage decreases from **$`r round(env_cost,2)`** to **$`r round(env_cost_tax,2)`,** a change of **$`r round(env_change, 2)`.**

g) The total tax revenue generated from the policy was **$`r round(tax_rev,2)`.**


### **5. Redistribution**

```{r}
price_c_51 <- 1.97
price_c_75 <- 2.89
price_c_100 <- 3.86
price_c_125 <- 4.82
price_c_150 <- 5.78

#total share of demand for $51/ton tax
demand_high_og <- 415576.92
demand_low_og <- 121545.45

low_share <- (demand_low_og/(demand_high_og + demand_low_og))*100
high_share <- (demand_high_og/(demand_high_og+demand_low_og))*100

```

estimates of distributional effects of a carbon tax with different SCC values are as follows: 

**$51/ton:**
```{r}
low_share_51 <- (low_share * tax_rev)/100
high_share_51 <- (high_share * tax_rev)/100


```

tax is 51/(2204/0.85) = **$0.0197/kWh**

  - welfare effects on "high" income consumers: 
  
  - welfare effects on "low" income consumers: 
  
  - welfare effects on electricity producers:
  
**$75/ton:**

tax is 75/(2204/0.85) = **$0.0289/kWh**



  - welfare effects on "high" income consumers: 
  
  - welfare effects on "low" income consumers: 
  
  - welfare effects on electricity producers:

**$100/ton:**

tax is 100/(2204/0.85) = **$0.0386/kWh**

  - welfare effects on "high" income consumers: 
  
  - welfare effects on "low" income consumers: 
  
  - welfare effects on electricity producers:
  
**$125/ton:**

tax is 125/(2204/0.85) = **$0.0482/kWh**

  - welfare effects on "high" income consumers: 
  
  - welfare effects on "low" income consumers: 
  
  - welfareeffects on electricity producers:
  
**$150/ton:**

tax is 150/(2204/0.85) = **$0.0578/kWh**

  - welfare effects on "high" income consumers: 
  
  - welfare effects on "low" income consumers: 
  
  - welfare effects on electricity producers:
  

### **6. Solar PV**

```{r}
# Demand model from sample code

demand_solar <- function(p, model){
  q <- (p - 0.5*model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
```


```{r}
# Aggregate demand from code sample 

demand_agg_solar <- function(p){
  q <- demand_solar(p, demand_low) + demand_solar(p, demand_high)
  return(q)
}


```

```{r}
# Make a vector and extract the lm 
price_solar = seq(0, 30, length.out = 100)

Qagg_solar <- map(price_solar, demand_agg_solar) %>% 
  unlist()

agg_df_solar<- tibble(Qagg_solar = Qagg_solar, price_solar = price_solar)

demand_agg_eq_solar <- lm(price_solar ~ Qagg_solar, data = agg_df_solar) 
```

```{r}
# Switch the equations above so Q is on one side and then add

agg_slope_solar <- demand_agg_eq_solar$coefficients[2]
agg_int_solar <- demand_agg_eq_solar$coefficients[1]

#agg_slope_solar
#agg_int_solar

solar_q <- 242802.548

solar_env_cost <- (solar_q * 1.97)/100

new_tax_6 <- (1.97/(solar_env_cost/env_cost_tax))

```

new aggregate demand function is P = 21.032 - 0.0000593Q.
supply curve: P = 0.0000192Q + 1.97

a) total electricity consumption: **`r round(solar_q, 2)` kWh**

b) total environmental externality: **$`r round(solar_env_cost,2)`.**

c) a **`r round (new_tax_6,2)` cent** tax on electricity would make the total environmental damage in the normal scenario ($`r round (env_cost,2)`) equal the new, reduced environmental damages from the demand shift driven by the switch to solar. 

